Leadership across cultures and contexts: an exploratory analysis

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Edits to discussion section. Mediate -> Strategize.

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)). Some common leadership dimensions also did not vary by the four factors we examined: group context, continental region, subsistence strategy, and leader sex. These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors. See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), organizing collective actions, and

@@

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders [but see @garfield_investigating_2019]. These dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.

@@

Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019]. But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, which was below our arbitrary 60% threshold for "universality". See Figure \@ref(fig:varsupportcb).

>>>>

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)). The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial. See Figure \@ref(fig:featuressupportplot).

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, continental region, subsistence strategy, and leader sex. These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors. See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), strategizing, and organizing collective actions.

@@

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.

@@

Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019]. But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality". See Figure \@ref(fig:varsupportcb).
